---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# acronymr

acronymr is an R package to facilitate the use of acronyms in rmarkdown documents. Acronyms and their definitions are stored in a data frame which are then referenced inline to place the formatted acronym in the rendered document. Initial use of an acronym will define the acronym while subsequent use will just place the acronym. Options are available to use the plural form, an alternate form which may include special formatting such as italics or super and subscripts, or the definition.

acronymr tracks which acronyms are used and provides a function to return a data frame of used acronyms suitable for output in a table.

## Installation

You can install acronymr from github with:

```{r gh-installation, eval = FALSE, echo=TRUE}
# install.packages("devtools")
devtools::install_github("HJAllen/acronymr")
```

## Example


```{r acronyms, echo=TRUE, results='markup'}
library(acronymr)

acronyms <-
  # Create acronyms lookup table
  tibble::tribble(
    ~Acronym, ~Alternate, ~Definition,
    "GMO", NA, "Genetically Modified Organism",
    'IMO', NA,'In My Opinion',
    'PCR', NA, 'Polymerase Chain Reaction',
    'DNA', NA, 'Deoxyribonucleic Acid',
    'H2NO3', 'H~2~NO~3~', 'Nitric Acid')
# Add inclusion flag
acronyms$Include <- FALSE
# Alphabetize acronyms for ease of reading
acronyms <- acronyms[order(acronyms$Acronym),]

acronyms
```

When an acronym is used, if Include is set to `FALSE` indicating first usage, the return value is the long form of the acronym followed by either the acronym name or alternate form depending on `altFrm`. Plural forms can be generated using the parameter `p`. Subsequent uses of the acronym will return the acronym formatted as requested.

The following:

The `` `r
ac("DNA")` `` of a `` `r
ac("GMO")` `` has been modified. `` `r
ac("PCR")` `` can be used to characterize `` `r
ac("DNA")` `` in `` `r
ac("GMO", p = TRUE)` ``. Be careful to avoid introduction of `` `r
ac("H2NO3", altFrm = TRUE)` ``.

Produces:

The `r ac("DNA")` of a `r ac("GMO")` has been modified. `r ac("PCR")` can be used to characterize `r ac("DNA")` in `r ac("GMO", p = TRUE)`. Be careful to avoid introduction of `r ac("H2NO3", altFrm = TRUE)`.

`acronyms$Include` is updated to `TRUE`.

A data frame containing used acronyms:
```{r ChunkToCacheAcronyms, results='markup'}
acTable("acronyms")
```
